---
output: github_document
always_allow_html: true
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# MargIntTmle

<!-- badges: start -->
<!-- badges: end -->

The goal of MargIntTmle is to estimate marginal interaction effects using g-computation, IPTW or TMLE. Interaction effects are calculated from the parameters of a Marginal Structural Model (MSM) estimated using the [`ltmle`](https://github.com/joshuaschwab/ltmle) R package.

## Installation

You can install the development version of MargIntTmle from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("benoitlepage/MargIntTmle")

# in case of error message "ERROR: loading failed for 'i386'"
# use the following command instead
# (which should force it to only build the package for your currently running R version.):
# devtools::install_github("benoitlepage/MargIntTmle", INSTALL_opts=c("--no-multiarch"))
```

## First example, with a binary outcome

In this first example, we simulate a data set of `N` = 1000 rows, with three baseline confounders (`conf1`, `conf2`, and `conf3`), two exposures (`sex` and `env`) and one outcome `hlth.outcome`. We used the default parameters defined in the `param.causal.model()` function.

```{r simulate_data}
require(MargIntTmle)
set.seed(12345)
df <- generate.data(N = 1000, b = param.causal.model())
head(df)
```

The `int.ltmleMSM` function is used to call the `ltmleMSM` function from the `ltmle` package, in order to estimate marginal interaction effects from a Marginal structural model. In this example, we use the TMLE estimator.

Several quantities of interest are then calculated using the `estim.int.effects` function.

```{r example_binary, message=FALSE, warning=FALSE}
require(ltmle)
require(SuperLearner)
# define Q and g formulas following the argument notation of the ltmle package:
Q_formulas = c(hlth.outcome="Q.kplus1 ~ conf1 + conf2 + conf3 + sex * env")
g_formulas = c("sex ~ conf1 + conf2","env ~ conf1 + conf3")
# choose a set of fitting libraries to pass to SuperLearner:
SL.library = list(Q=list("SL.glm"),g=list("SL.glm"))
# apply the int.ltmleMSM function. In order to apply the TMLE and IPTW estimators, 
# gcomp argument is set to FALSE.
interaction.ltmle <- int.ltmleMSM(data = df,
                                  Qform = Q_formulas,
                                  gform = g_formulas,
                                  Anodes = c("sex", "env"),
                                  Lnodes = c("conf1", "conf2", "conf3"),
                                  Ynodes = c("hlth.outcome"),
                                  SL.library = SL.library,
                                  gcomp = FALSE,
                                  iptw.only = FALSE,
                                  survivalOutcome = FALSE,
                                  variance.method = "ic")
# several quantities of interest for interaction effects are calculated using the 
# estim.int.effects() function
est.tmle <- estim.int.effects(interaction.ltmle, estimator = "tmle")
```

The results can be presented in a table following [`Knol and VanderWeele`](https://doi-org.proxy.insermbiblio.inist.fr/10.1093/ije/dyr218) recommendations (2012). The out.table object contains the table and the interaction.effects object contains the additive, multiplicative interaction effects and the RERI.

The interaction table can be rendered using the `kableExtra` package.
```{r table_interaction}
table_inter <- out.int.table(int.r = est.tmle)
table_inter$out.table
table_inter$interaction.effects
library(kableExtra)
kbl(table_inter$out.table,
    caption = "Interaction effects estimated by TMLE") %>%
  kable_classic() %>%
  footnote(general = table_inter$interaction.effects)
```

We can also plot the results using the `out.int.fig()` function from the output of the `estim.int.effects()` function.

```{r interaction_plot}
out.int.fig(est.tmle)
```


## Second example, with a continuous outcome
```{r simulate_data_continuous}
# require(MargIntTmle)
set.seed(54321)
beta <- param.causal.model(Y_type = "continuous", b_Y = 100, b_L1_Y = 10, b_L2_Y = 30,
                           b_L3_Y = -20, b_A1_Y = 10, b_A2_Y = 30, b_A1A2_Y = 20,
                           se_Y = 20)
df.cont.Y <- generate.data(N = 1000, b = beta, Y_type = "continuous")
summary(df.cont.Y)
```

When using the `int.ltmleMSM`, the `ltmleMSM` function from the `ltmle` package transforms the outcome on a 0 to 1 scale: 

$Z = \frac{Y - \min(Y)}{\max(Y) - \min(Y)}$

The Marginal structural model estimated by the `ltmle` package is a quasi-binomial regression of $Z$ conditional on the exposures $A_1$ and $A_2$

$\text{logit} Z = \beta_0 + \beta_{A_1} A_1 + \beta_{A_2} A_2 + \beta_{A_1 \ast A_2} (A_1 \ast A_2)$


```{r example_continuous, message=FALSE, warning=FALSE}
require(ltmle)
require(SuperLearner)
# define Q and g formulas following the argument notation of the ltmle package:
Q_formulas = c(hlth.outcome="Q.kplus1 ~ conf1 + conf2 + conf3 + sex * env")
g_formulas = c("sex ~ conf1 + conf2","env ~ conf1 + conf3")
# choose a set of fitting libraries to pass to SuperLearner:
SL.library = list(Q=list("SL.glm"),g=list("SL.glm"))
# apply the int.ltmleMSM function. In order to apply the TMLE and IPTW estimators, 
# gcomp argument is set to FALSE.
continuous.interaction <- int.ltmleMSM(data = df.cont.Y,
                                       Qform = Q_formulas,
                                       gform = g_formulas,
                                       Anodes = c("sex", "env"),
                                       Lnodes = c("conf1", "conf2", "conf3"),
                                       Ynodes = c("hlth.outcome"),
                                       SL.library = SL.library,
                                       gcomp = FALSE,
                                       iptw.only = FALSE,
                                       survivalOutcome = FALSE,
                                       variance.method = "ic")

# We can see that the outcome has been transformed. 
# The range used to transform the outcome is given in the output
continuous.interaction$ltmle_MSM$transformOutcome
attr(continuous.interaction$ltmle_MSM$transformOutcome, "Yrange")

# The parameters of the MSM are estimated on the logit scale
continuous.interaction$ltmle_MSM$msm$family
# where the coefficients beta_0, beta_A1, beta_A2, and beta_A1_A2 are respectively
continuous.interaction$ltmle_MSM$msm$coefficients

## OLD TO UPDATE
# several quantities of interest for interaction effects are calculated using the 
# estim.int.effects() function
# est.tmle <- estim.int.effects(interaction.ltmle, estimator = "tmle")
```
