---
title: "Test performance on simulations"
author: "Benoit Lepage"
date: "2023-05-15"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(magrittr)
```

# Simulations

Simulation setting

```{r load_results, include=FALSE}
perf_gcomp_bin <- readRDS(file = "perf_gcomp_bin")
perf_iptw_bin <- readRDS(file = "perf_iptw_bin")
perf_tmle_bin <- readRDS(file = "perf_tmle_bin")

perf_gcomp_bin$bias <- perf_gcomp_bin$bias * 100
perf_gcomp_bin$std.bias <- perf_gcomp_bin$std.bias * 100
perf_gcomp_bin$av.estim.se <- perf_gcomp_bin$av.estim.se * 100
perf_gcomp_bin$cov <- perf_gcomp_bin$cov * 100

perf_iptw_bin$bias <- perf_iptw_bin$bias * 100
perf_iptw_bin$std.bias <- perf_iptw_bin$std.bias * 100
perf_iptw_bin$av.estim.se <- perf_iptw_bin$av.estim.se * 100
perf_iptw_bin$cov <- perf_iptw_bin$cov * 100

perf_tmle_bin$bias <- perf_tmle_bin$bias * 100
perf_tmle_bin$std.bias <- perf_tmle_bin$std.bias * 100
perf_tmle_bin$av.estim.se <- perf_tmle_bin$av.estim.se * 100
perf_tmle_bin$cov <- perf_tmle_bin$cov * 100

perf_gcomp_cont <- readRDS(file = "perf_gcomp_cont")
perf_iptw_cont <- readRDS(file = "perf_iptw_cont")
perf_tmle_cont <- readRDS(file = "perf_tmle_cont")

perf_gcomp_cont$std.bias <- perf_gcomp_cont$std.bias * 100
perf_gcomp_cont$cov <- perf_gcomp_cont$cov * 100

perf_iptw_cont$std.bias <- perf_iptw_cont$std.bias * 100
perf_iptw_cont$cov <- perf_iptw_cont$cov * 100

perf_tmle_cont$std.bias <- perf_tmle_cont$std.bias * 100
perf_tmle_cont$cov <- perf_tmle_cont$cov * 100
```

# 2) Results for the binary outcome
## 2.1) Estimation by g-computation

```{r table_gcomp_bin, echo=FALSE}
perf_gcomp_bin %>%
  kable(format = "latex",
        digits = c(3,5,2,5,2,1), 
        col.names = c("Bias (%)",
                      "Variance",
                     "Std.bias (%)",
                     "MSE",
                     "Av.SE (%)", 
                     "Coverage (%)"),
        caption = "Results from g-computation estimation with a binary outcome",
        escape = TRUE)
```
Note that the standard error was not estimated in 1 simulation out of 1000, resulting in a missing average standard error and coverage for the $\log\text{RERI}$ indicator.

## 2.2) Estimation by IPTW
```{r table_iptw_bin, echo=FALSE}
perf_iptw_bin %>%
  kable(format = "latex",
        digits = c(3,5,2,5,2,1), 
        col.names = c("Bias (%)",
                      "Variance",
                     "Std.bias (%)",
                     "MSE",
                     "Av.SE (%)", 
                     "Coverage (%)"),
        caption = "Results from IPTW estimation with a binary outcome",
        escape = TRUE)
```

## 2.3) Estimation by TMLE
```{r table_tmle_bin, echo=FALSE}
perf_tmle_bin %>%
  kable(format = "latex",
        digits = c(3,5,2,5,2,1), 
        col.names = c("Bias (%)",
                      "Variance",
                     "Std.bias (%)",
                     "MSE",
                     "Av.SE (%)", 
                     "Coverage (%)"),
        caption = "Results from TMLE estimation with a binary outcome",
        escape = TRUE)
```


# 3) Results for the continuous outcome
## 3.1) Estimation by g-computation

```{r table_gcomp_cont, echo=FALSE}
perf_gcomp_cont %>%
  kable(format = "latex",
        digits = c(3,2,2,2,1,1), 
        col.names = c("Bias (%)",
                      "Variance",
                     "Std.bias (%)",
                     "MSE",
                     "Av.SE (%)", 
                     "Coverage (%)"),
        caption = "Results from g-computation estimation with a continuous outcome",
        escape = TRUE)
```

## 3.2) Estimation by IPTW

```{r table_iptw_cont, echo=FALSE}
perf_iptw_cont %>%
  kable(format = "latex",
        digits = c(3,2,2,2,1,1), 
        col.names = c("Bias (%)",
                      "Variance",
                     "Std.bias (%)",
                     "MSE",
                     "Av.SE (%)", 
                     "Coverage (%)"),
        caption = "Results from IPTW estimation with a continuous outcome",
        escape = TRUE)
```

## 3.3) Estimation by TMLE

```{r table_tmle_cont, echo=FALSE}
perf_tmle_cont %>%
  kable(format = "latex",
        digits = c(3,2,2,2,1,1), 
        col.names = c("Bias (%)",
                      "Variance",
                     "Std.bias (%)",
                     "MSE",
                     "Av.SE (%)", 
                     "Coverage (%)"),
        caption = "Results from TMLE estimation with a continuous outcome",
        escape = TRUE)
```

